Multiple Output Regression

### 2014-02-18 paper/paper.tex
fix equations

\noindent Also as a consequence of sufficient statistics, the approximate prediction for test targets $\vfstar$ is
\begin{align}
\nonumber
p(\vfstar | \y, \X_*) &= \int p(\vfstar | \f, \u, \X_*) q(\f, \u | \y) \der \f \der \u \\
\nonumber
&= \int p(\vfstar | \u) q(\u| \y) p(\f | \u) \der \f \der \u \\
\label{eq:sorprediction}
&= \int p(\vfstar | \u) q(\u| \y) \der \u
\end{align}
which is the same as the prediction of a SoR model (of course with a different posterior in $q(\u | \y)$).

@@

\noindent Each individual term $l_{in} = \llangle \log p(y_{in} | g_n, h_{in}) \rrangle_{p(\g,\h_i | \ug, \u_i)}$ can be computed as follows:
\begin{align}
\nonumber
l_{in} &= \int \log p(y_{in} | g_n, h_{in}) p(\g | \ug) p(\h_i | \u_i) \der \g \der \h_i \\
\nonumber
&= \int \log \Normal(y_{in} ; w_i g_n + h_{in}, \beta_i^{-1}) 
\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
\Normal(h_{in} ; \mu_{in}, (\tilde{k}_{inn}) \der g_n \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int (y_{in} - w_i g_n - h_{in}) \beta (y_{in} - w_i g_n - h_{in})
\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der g_n \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int \left[(y_{in} - h_{in} - w_i \mu_{gn}) \beta (y_{in} - h_{in} - w_i \mu_{gn}) + w_i^2 \beta \tilde{k}_{gnn} \right] 
\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} w_i^2 \beta \tilde{k}_{gnn}
- \frac{1}{2} \beta \tilde{k}_{inn} - \frac{1}{2} (y_{in} - w_i \mu_{gn} - \mu_{in}) \beta (y_{in} - w_i \mu_{gn} - \mu_{in}) \\
&= \log \Normal(y_{in}; w_i \mu_{gn} + \mu_{in}, \beta_i^{-1})  - \frac{1}{2} \beta_i (w_i^2 \tilde{k}_{gnn} + \tilde{k}_{inn}),
\end{align}
where $\tilde{k}_{gnn} = (\tilde{\K}_g)_{n,n}, \tilde{k}_{inn} = (\tilde{\K}_i)_{n,n}, \mu_{gn} = (\Mu_g)_n, \text{ and } \mu_{in} = (\Mu_i)_n$.
Notice that the above expression is very similar to the single-output case.
Hence our derivation here can be seen as the \textit{generalization} of SVI from the standard GP regression to multiple-output regression.

\noindent Substituting $l_{in}$ into equation \ref{eq:elbo} and carrying similar integration we get,
\begin{align}
\nonumber
\log p(\y)
\ge& \sum_{i=1}^P \sum_{n=1}^N
 \left( \log \Normal(y_{in}; \tilde{\mu}_{in}, \beta^{-1})
         - \frac{1}{2} \beta_i (w_i^2 \tilde{k}_{gnn} + \tilde{k}_{inn})
         - \frac{1}{2} \trace \left( w_i^2 \S_g \mat{\Lambda}_{gn} + \S_i \mat{\Lambda}_{in} \right)
\right) \\
&- \left( \KL[q(\ug) || p(\ug)] + \sum_{i=1}^P \KL[q(\u_i) || p(\u_i)] \right) \define \calL,
\end{align}
where 
\begin{align}
\tilde{\mu}_{in}
&= w_i k_g(\x_n, \Zg)k_g(\Zg,\Zg)^{-1}\m_g + k_i(\x_n, \Zi)k_i(\Zi,\Zi)^{-1}\m_i \\
\mat{\Lambda}_{gn}
&= \beta k_g(\Zg,\Zg)^{-1} k_g(\Zg, \x_n) k_g(\x_n, \Zg) k_g(\Zg,\Zg)^{-1} \\
\mat{\Lambda}_{in}
&= \beta k_i(\Zi,\Zi)^{-1} k_i(\Zi, \x_n) k_i(\x_n, \Zg) k_i(\Zg,\Zg)^{-1}.
\end{align}

@@

\noindent As a function of the parameters of $q(\ug)$, the lowerbound $\calL$ is:
\begin{align}
\nonumber
\calL_g \define&
 \sum_{i=1}^P \left( \log \Normal(\y_i; w_i \A_g(\oi,:) \m_g + \A_i \m_i, \beta_i^{-1} \I)
 - \frac{1}{2} \beta_i \trace w_i^2 \tilde{\K}_g(\oi,\oi)
 - \frac{1}{2} \beta_i \trace w_i^2 \S_g \A_g^T \A_g 
  \right) \\
  &- \frac{1}{2} \log |k_g(\Zg,\Zg) \S_g^{-1}| -\frac{1}{2} \trace k_g(\Zg,\Zg)^{-1} (\m_g \m_g^T + \S_g) ,
\end{align}
where $\A_g = k_g(\X,\Z_g)k_g(\Z_g,\Z_g)^{-1}$, which gives $\A_g(\oi,:) = k_g(\X_i,\Zg) k_g(\Z_g,\Z_g)^{-1}$, and  
$\A_i = k_i(\X_i,\Z_i)k_i(\Z_i,\Z_i)^{-1}$. \\

@@

\noindent The derivatives of $\calL_i$ wrt $\m_i$ and $\S_i$ are given by:
\begin{align}
\deriv{\calL_g}{\m_i}
& = \beta_i \A_i^T (\y_i - w_i \A_g(\oi,:)^T \m_g) - \left[k_i(\Zi,\Zi)^{-1} +  \beta_i \A_i^T \A_i \right] \m_i \\
\deriv{\calL_g}{\S_g} 
&= \frac{1}{2} \S_i^{-1} - \frac{1}{2} \left[ k_i(\Zi,\Zi)^{-1} + \beta_i \A_i^T \A_i \right].
\end{align}

@@

%\subsubsection{Update Equations}
% and comment on the intuition of the update equations: e.g. y(x) - g(x) for h(x) 
\subsection{Prediction}
The predictive distribution of the $i$-th output for a test input $\x_*$ is 
\begin{align}
p(\fstar | \y, \x_*) = \int \Normal(\fstar; w_i g_* + h_{i*}, 0) p(g_* | \y, \x_*) p(h_{i*} | \y, \x_*) \der g_* \der h_{i*},
\end{align}
where $p(g_* | \y, \x_*) = \Normal(g_*; \mu_{g*}, s_{g*})$ and $p(h_{i*} | \y, \x_*) = p(h_{i*}; \mu_{i*}, s_{i*})$ are the predictive distributions of the sparse GPs as given in \ref{eq:sorprediction}.
Therefore we have:
\begin{align}
p(\fstar | \y, \x_*) = \Normal(\fstar; w_i \mu_{g*} + \mu_{i*}, s_{g*} + s_{i*}). 
\end{align}

>>>>

\noindent Also as a consequence of sufficient statistics, the approximate prediction for test targets $\vfstar$ at test inputs $\X_*$ is
\begin{align}
\nonumber
p(\vfstar | \y, \X_*) &= \int p(\vfstar | \f, \u, \X_*) q(\f, \u | \y) \der \f \der \u \\
\nonumber
&= \int p(\vfstar | \u) q(\u| \y) p(\f | \u) \der \f \der \u \\ \nonumber
&= \int p(\vfstar | \u) q(\u| \y) \der \u \\
\label{eq:sorprediction}
&= \Normal(\vfstar; \bs{\mu_*},\vec{s_*})
\end{align}
where,
\begin{align}
\nonumber
\bs{\mu_*} &= \K_{*z} \K_{zz}^{-1}\m \\ 
\nonumber
\S_* &= \K_{**} - \K_{*z} \left(\K_{zz}^{-1} - \K_{zz}^{-1} \S \K_{zz}^{-1} \right) \K_{*z}^T.
\end{align}
 Here $\K_{*z}$ is the covariance matrix between test and inducing inputs, $\K_{zz}$ is the covariance matrix of the inducing inputs.

@@

\noindent Each individual term $l_{in} = \llangle \log p(y_{in} | g_n, h_{in}) \rrangle_{p(\g,\h_i | \ug, \u_i)}$ can be computed as follows:
\begin{align}
\nonumber
l_{in} &= \int \log p(y_{in} | g_n, h_{in}) p(\g | \ug) p(\h_i | \u_i) \der \g \der \h_i \\
\nonumber
&= \int \log \Normal(y_{in} ; w_i g_n + h_{in}, \beta_i^{-1}) 
\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
\Normal(h_{in} ; \mu_{in}, (\tilde{k}_{inn}) \der g_n \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int (y_{in} - w_i g_n - h_{in}) \beta_i (y_{in} - w_i g_n - h_{in})
\Normal(g_n ; \mu_{gn}, \tilde{k}_{gnn})
\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der g_n \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} \int \left[(y_{in} - h_{in} - w_i \mu_{gn}) \beta_i (y_{in} - h_{in} - w_i \mu_{gn}) + w_i^2 \beta_i \tilde{k}_{gnn} \right] 
\Normal(h_{in} ; \mu_{in}, \tilde{k}_{inn}) \der h_{in} \\
\nonumber
&= -\frac{1}{2} \log 2 \pi \beta_i^{-1} - \frac{1}{2} w_i^2 \beta_i \tilde{k}_{gnn}
- \frac{1}{2} \beta_i \tilde{k}_{inn} - \frac{1}{2} (y_{in} - w_i \mu_{gn} - \mu_{in}) \beta_i (y_{in} - w_i \mu_{gn} - \mu_{in}) \\
&= \log \Normal(y_{in}; w_i \mu_{gn} + \mu_{in}, \beta_i^{-1})  - \frac{1}{2} \beta_i (w_i^2 \tilde{k}_{gnn} + \tilde{k}_{inn}),
\end{align}
where $\tilde{k}_{gnn} = (\tilde{\K}_g)_{n,n}, \tilde{k}_{inn} = (\tilde{\K}_i)_{n,n}, \mu_{gn} = (\Mu_g)_n, \text{ and } \mu_{in} = (\Mu_i)_n$.
Notice that the above expression is very similar to the single-output case.
Hence our derivation here can be seen as the \textit{generalization} of SVI from the standard GP regression to multiple-output regression.

\noindent Substituting $l_{in}$ into equation \ref{eq:elbo} and carrying similar integration we get,
\begin{align}
\nonumber
\log p(\y)
\ge& \sum_{i=1}^P \sum_{n=1}^N
 \left( \log \Normal(y_{in}; \tilde{\mu}_{in}, \beta^{-1})
         - \frac{1}{2} \beta_i (w_i^2 \tilde{k}_{gnn} + \tilde{k}_{inn})
         - \frac{1}{2} \trace \left( w_i^2 \S_g \mat{\Lambda}_{gn} + \S_i \mat{\Lambda}_{in} \right)
\right) \\
&- \left( \KL[q(\ug) || p(\ug)] + \sum_{i=1}^P \KL[q(\u_i) || p(\u_i)] \right) \define \calL,
\end{align}
where 
%TODO \Lambda_gn depends on the output i so need a better notation
\begin{align}
\tilde{\mu}_{in}
&= w_i k_g(\x_n, \Zg)k_g(\Zg,\Zg)^{-1}\m_g + k_i(\x_n, \Zi)k_i(\Zi,\Zi)^{-1}\m_i \\
\mat{\Lambda}_{gn}
&= \beta_i k_g(\Zg,\Zg)^{-1} k_g(\Zg, \x_n) k_g(\x_n, \Zg) k_g(\Zg,\Zg)^{-1} \\
\mat{\Lambda}_{in}
&= \beta_i k_i(\Zi,\Zi)^{-1} k_i(\Zi, \x_n) k_i(\x_n, \Zg) k_i(\Zg,\Zg)^{-1}.
\end{align}

@@

\noindent As a function of the parameters of $q(\ug)$, the lowerbound $\calL$ is:
\begin{align}
\nonumber
\calL_g \define&
 \sum_{i=1}^P \log \Normal(\y_i; w_i \A_g(\oi,:) \m_g + \A_i \m_i, \beta_i^{-1} \I)  \\
 &- \frac{1}{2} \sum_{i=1}^P \bigg(\beta_i \trace w_i^2 \tilde{\K}_g(\oi,\oi) 
 + \beta_i \trace w_i^2 \S_g \A_g(\oi,:)^T \A_g(\oi,:) \bigg)
 \\
 &- \frac{1}{2} \log |k_g(\Zg,\Zg) \S_g^{-1}| -\frac{1}{2} \trace k_g(\Zg,\Zg)^{-1} (\m_g \m_g^T + \S_g) ,
\end{align}
where $\A_g = k_g(\X,\Z_g)k_g(\Z_g,\Z_g)^{-1}$, which gives $\A_g(\oi,:) = k_g(\X_i,\Zg) k_g(\Z_g,\Z_g)^{-1}$, and  
$\A_i = k_i(\X_i,\Z_i)k_i(\Z_i,\Z_i)^{-1}$. \\

@@

\noindent The derivatives of $\calL_i$ wrt $\m_i$ and $\S_i$ are given by:
\begin{align}
\deriv{\calL_i}{\m_i}
& = \beta_i \A_i^T (\y_i - w_i \A_g(\oi,:)^T \m_g) - \left[k_i(\Zi,\Zi)^{-1} +  \beta_i \A_i^T \A_i \right] \m_i \\
\deriv{\calL_i}{\S_i} 
&= \frac{1}{2} \S_i^{-1} - \frac{1}{2} \left[ k_i(\Zi,\Zi)^{-1} + \beta_i \A_i^T \A_i \right].
\end{align}

@@

%\subsubsection{Update Equations}
% and comment on the intuition of the update equations: e.g. y(x) - g(x) for h(x) 
\subsection{Prediction}
The predictive distribution of the $i$-th output for a test input $\x_*$ is 
\begin{align}
p(\fstar | \y, \x_*) = \int \Normal(\fstar; w_i g_* + h_{i*}, 0) p(g_* | \y, \x_*) p(h_{i*} | \y, \x_*) \der g_* \der h_{i*},
\end{align}
where $p(g_* | \y, \x_*) = \Normal(g_*; \mu_{g*}, s_{g*})$ and $p(h_{i*} | \y, \x_*) = p(h_{i*}; \mu_{i*}, s_{i*})$ are the predictive distributions of the sparse GPs as given in \ref{eq:sorprediction}.
Therefore we have:
\begin{align}
p(\fstar | \y, \x_*) = \Normal(\fstar; w_i \mu_{g*} + \mu_{i*}, w_i^2 s_{g*} + s_{i*}). 
\end{align}
